---
title: "01_Dataset_Creation"
author: "Jay Chandra and Anushka Bhaskar"
date: "3/29/2022"
output: pdf_document
---

```{r}
library(dplyr)
library(tseries)
library(forecast)
library(ggplot2)
library(lubridate)
library(tscount)
library(gridExtra)
library(usmap) 
library(ggthemes)
library(readxl)
library(stringr)
library(pracma)
library(maps)
library(mapdata)
```


# Read in Datasets
```{r}
country_1 <- read.delim("data/Year.txt")
country_2 <- read.delim("data/Year_2.txt")
state_1 <- read.delim("data/State_Month.txt")
state_2 <- read.delim("data/State_Month_2.txt")
county_1 <- read.delim("data/County_Year.txt")
county_2 <- read.delim("data/County_Year_2.txt")
state_pop <- read.csv('data/State_Population.csv')
colnames(state_pop) <- c('State', 'Population', 'Drop1', 'Drop2')
state_pop = state_pop %>% select('State', 'Population')
state_pop$Population = as.numeric(gsub(",","",state_pop$Population))
cnames = c("County", "Pop", "Pop_2", "Pop_3")
county_pop <- read_excel('data/County_Population.xlsx', col_names = cnames, skip = 3)
county_pop = na.omit(county_pop)
county_pop[ , 'fips'] <- rep("NA", nrow(county_pop))
```

```{r}
# COVID-19 Data
cases = read.csv('data/cases.csv')
cases = subset (cases, select = -c(UID, iso2, iso3, code3, FIPS, Admin2, Country_Region, Lat, Long_, Combined_Key))
cases_state <- cases %>% group_by(Province_State) %>% summarise(across(everything(), sum))
```

```{r}
# Edits COVID-19 Dataframe so that it only includes the data at monthly level
cases_state_month <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = nrow(cases_state),
                          ncol = 0))
cases_state_month["state"] = cases_state["Province_State"]
indexM = vector()
startMonthP = "1."
for (i in seq(2, ncol(cases_state))){
    startMonth = substring(colnames(cases_state)[i], 2, 3)
    if (startMonth == startMonthP){
      indexM = c(indexM, i)
    }
    else{
      newData = cases_state[, indexM]
      newCol = newData[, ncol(newData)] - newData[, 1]
     cases_state_month[paste0(substring(colnames(cases_state)[i-1], 2, 3), substring(colnames(cases_state)[i-1], nchar(colnames(cases_state)[i-1]) - 1, nchar(colnames(cases_state)[i-1])))] = newCol
      startMonth = substring(colnames(cases_state)[i], 2, 3)
      startMonthP = startMonth
      indexM = c(i)
    }
}
cases_state_month = cases_state_month[, c(1, 4:(ncol(cases_state_month)-7))]
```


```{r}
# finds fips code from county names
for (row in 1:nrow(county_pop)) {
  if (row %% 100 == 0){
    print(row)
  }
    name <- county_pop[row, "County"]
    name = name$County
    name <- substr(name, 2, nchar(name))
    name <- strsplit(name, ', ', fixed=T)
    tryfips <- tryCatch({
        fips(state = name[[1]][2], county = name[[1]][1])
    },
    error = function(e){
      NA
    })
    county_pop[row, 'fips'] <- tryfips
}
```

```{r}
state_1 = state_1[state_1$Notes == "", ]
state_2 = state_2[state_2$Notes == "" & state_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
colnames(state_2)[2] <- 'State'
colnames(state_2)[3] <- 'State.Code'
state = rbind(state_1, state_2)
```

```{r}
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

# sARIMA Modelling Drug-Related National Level

```{r}
# Forecasting at the national level drug-related
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                    seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
clrs = c("blueviolet", "chartreuse3", "chartreuse3", "blueviolet")
autoplot(t_Country, series="Data", size = 0.7) + 
  geom_vline(xintercept = 2020.12, linetype = "dashed", color = "black", size = 0.8)+
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Pandemic") + 
  xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "none") + scale_x_continuous(breaks=c(2015, 2016, 2017, 2018, 2019, 2020, 2021)) + 
scale_color_manual(values=clrs)
ggsave(paste('plots/',"US States Time"),
  device = "png", width = 6, height = 5, dpi = 600)
```

```{r}
# Population normalization
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

# sARIMA Modelling Drug-Related State Level, Correlation with COVID-19 cases

```{r}
plots = list()
correlation_cases = vector()
state_names = as.character(unique(state$State))
df_params <- data.frame(matrix(nrow = 51, ncol = 7))
colnames(df_params) <- c('State', 'p', 'd', 'q', 'P', 'D', 'Q')
df_excess <- data.frame(matrix(nrow = 51, ncol = 8))
colnames(df_excess) <- c('State', 'Excess', 'Lower95', 'Upper95', 'Excess_2', 'Lower95_2', 'Upper95_2', 'correlation')
df_excess_time <- data.frame(matrix(nrow = 51, ncol = 19))
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  cases = cases_state_month[cases_state_month$state == state_name, 2:ncol(cases_state_month)]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_State = ts(State_data$Deaths[1:(length(State_data$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)
  
  parameters = expand.grid(p = seq(0, 1), d = seq(0, 1), q = seq(0, 1),
              P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))
  
  best_aic = 10000
  for (j in seq(1, nrow(parameters))){
    
    fit <- tryCatch(
    {
        Arima(t_State, order = c(parameters$p[j], parameters$d[j], parameters$q[j]),
                      seasonal = c(parameters$P[j], parameters$D[j], parameters$Q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_State, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        Arima(t_State, order = c(1, 0, 0),
                      seasonal = c(0, 0, 1), method="ML")
        
        })
    })
      
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }

  f_fit<-forecast(best_fit, robust = TRUE, h = 18, level = 95)
  
  t_State2 = ts(State_data$Deaths[(length(State_data$Deaths) - 17):length(State_data$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)
  excess = as.numeric(t_State2) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(t_State2) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(t_State2) - as.numeric(f_fit[["upper"]])
  
  excess2 = as.numeric(t_State2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]
  lower952 = as.numeric(t_State2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]
  upper952 = as.numeric(t_State2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]
  
  df_excess_time[i, ] = c(state_name, excess)
  df_params[i, ] = c(state_name, best_params$p, best_params$d, best_params$q, best_params$P, best_params$D, best_params$Q)
  df_excess[i, ] = c(state_name, sum(excess), sum(lower95), sum(upper95), sum(excess2), sum(lower952), sum(upper952), cor(as.numeric(cases), as.numeric(excess)))
  
model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
 # g = autoplot(t_State, series="Data") + 
 #     autolayer(fit$fitted, series=model_name) +
 #     autolayer(f_fit, series="Prediction") +
 #     autolayer(t_State2, series = "True Post-Pandemic") +
 #     xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
 
 clrs = c("blueviolet", "chartreuse3", "chartreuse3", "blueviolet")
g = autoplot(t_State, series="Data", size = 0.7) + 
  geom_vline(xintercept = 2020.12, linetype = "dashed", color = "black", size = 0.8)+
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_State2, series = "True Pandemic") + 
  xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "none") + scale_x_continuous(breaks=c(2015, 2016, 2017, 2018, 2019, 2020, 2021)) + 
scale_color_manual(values=clrs)


 plots[[i]] = g
}

```

```{r}
# Save Plots
for (i in seq(1, length(plots))){
  ggsave(paste('plots/states/',state_names[i]),
  plot = plots[[i]],
  device = "png", width = 6, height = 5, dpi = 600)
}
```

```{r}
# Population Normalization
df_excess = merge(df_excess, state_pop)
df_excess = df_excess %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000), Excess_pop_2 = as.numeric(Excess_2)/(Population/100000), Lower95_pop_2 =as.numeric(Lower95_2)/(Population/100000), Upper95_pop_2 =as.numeric(Upper95_2)/(Population/100000))
colnames(df_excess)[1] = 'state'
```

```{r}
# Plot Map of US
plot_usmap(data = df_excess, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States"),
  device = "png")
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States"),
  device = "png")
```


```{r}
# Plot Map of US first couple of months
plot_usmap(data = df_excess, values = "Excess_pop_2", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States First 3 Months"),
  device = "png")
```

```{r}
# Plot all excess mortality on vertical plot
df_excess = df_excess[order(df_excess$Excess_pop, decreasing = T),]
df_excess$state = factor(df_excess$state, levels=df_excess[order(df_excess$Excess_pop), "state"])
df_excess = df_excess[complete.cases(df_excess), ]
vert_plot <- ggplot(
    df_excess,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot"),
  device = "png", width = 7, height = 9)
```

```{r}
# Plot all excess mortality on vertical plot first 3 months
df_excess = df_excess[order(df_excess$Excess_pop_2, decreasing = T),]
df_excess$state = factor(df_excess$state, levels=df_excess[order(df_excess$Excess_pop_2), "state"])
df_excess = df_excess[complete.cases(df_excess), ]
vert_plot <- ggplot(
    df_excess,
    aes(x = Excess_pop_2,xmin = Upper95_pop_2,xmax = Lower95_pop_2, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot First 3 Months"),
  device = "png", width = 7, height = 9)
```

# LOESS Approach (same organization as sARIMA section)

```{r}
country_arr = country$Deaths[1:(length(country$Deaths) - 18)]
country_arr2 = country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)]
loess_model <- loess(country_arr ~ as.numeric(1:length(country_arr)), control = loess.control(surface = "direct"))
fit_loess = predict(loess_model,se=TRUE)
predict_loess = predict(loess_model,as.numeric((length(country_arr) + 1):(length(country_arr) + 18)), se = TRUE)

x = linspace(2015, 2021.67, length(c(country_arr, country_arr2 )))

df_plots <- data.frame(matrix(nrow = length(country$Deaths), ncol = 4))
colnames(df_plots) <- c('x', 'mean', 'upper', 'lower')

df_plots$x <- x
df_plots$mean = c(fit_loess$fit, predict_loess$fit)
df_plots$lower = c((fit_loess$fit - 1.96*fit_loess$se.fit), (predict_loess$fit - 1.96*predict_loess$se.fit))
df_plots$upper = c((fit_loess$fit + 1.96*fit_loess$se.fit), (predict_loess$fit + 1.96*predict_loess$se.fit))
df_plots$actual = country$Deaths

ggplot(df_plots, aes(x=x,y=mean, ymin=lower, ymax=upper, color = 'LOESS model')) + 
geom_ribbon(color = NA,#border line color
                fill="#c1ef7d") + 
  geom_line(colour="#94cb00") + 
geom_line(aes(y=actual, color = 'Actual Data')) + 
geom_vline(xintercept = 2020.12, linetype = "dashed", color = "black", size = 0.8)+
theme_bw() +
xlab("Year") + ylab("Drug Overdose Deaths") + theme_bw() +
  theme(legend.position= "none", legend.title = element_blank()) +
  scale_color_manual(values = c("blueviolet", "chartreuse3")) +
  guides(color=guide_legend("Time Period")) + scale_x_continuous(breaks=c(2015, 2016, 2017, 2018, 2019, 2020, 2021))

ggsave(paste('plots/',"US States Time Loess"),
  device = "png", width = 6, height = 5, dpi = 600)
```

```{r}
# Pop Norm
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(country_arr2 - predict_loess$fit) / US_pop_norm
lower95 = sum(country_arr2 - df_plots$lower[(length(df_plots$lower) - 17):length(df_plots$lower)]) / US_pop_norm
upper95 = sum(country_arr2 - df_plots$upper[(length(df_plots$upper) -17):length(df_plots$upper)]) / US_pop_norm

excess2 = sum(country_arr2[1:4] - predict_loess$fit[1:4]) / US_pop_norm
lower952 = sum(country_arr2[1:4] - df_plots$lower[(length(df_plots$lower) - 17):(length(df_plots$lower)-14)]) / US_pop_norm
upper952 = sum(country_arr2[1:4] - df_plots$upper[(length(df_plots$upper) -17):(length(df_plots$upper)-14)]) / US_pop_norm
```

```{r}
# State Level
plots_loess = list()
state_names = as.character(unique(state$State))
df_excess_loess <- data.frame(matrix(nrow = 51, ncol = 7))
colnames(df_excess_loess) <- c('State', 'Excess', 'Lower95', 'Upper95', 'Excess_2', 'Lower95_2', 'Upper95_2')
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  if (mean(is.na(as.numeric(State_data$Deaths))) > 0.7){
    next
  }
  State_data$Deaths[is.na(State_data$Deaths)] <- 10
  
  state_arr = State_data$Deaths[1:(length(State_data$Deaths) - 18)]
  state_arr2 = State_data$Deaths[(length(State_data$Deaths) - 17):length(State_data$Deaths)]
  loess_model <- loess(state_arr ~ as.numeric(1:length(state_arr)), control = loess.control(surface = "direct"))
  fit_loess = predict(loess_model,se=TRUE)
  predict_loess = predict(loess_model,as.numeric((length(state_arr) + 1):(length(state_arr) + 18)), se = TRUE)
  
  x = linspace(2015, 2021.67, length(c(country_arr, country_arr2 )))

  df_plots <- data.frame(matrix(nrow = length(country$Deaths), ncol = 4))
  colnames(df_plots) <- c('x', 'mean', 'upper', 'lower')

  df_plots$x <- x
  df_plots$mean = c(fit_loess$fit, predict_loess$fit)
  df_plots$lower = c((fit_loess$fit - 1.96*fit_loess$se.fit), (predict_loess$fit - 1.96*predict_loess$se.fit))
  df_plots$upper = c((fit_loess$fit + 1.96*fit_loess$se.fit), (predict_loess$fit + 1.96*predict_loess$se.fit))
  df_plots$actual = State_data$Deaths
  excess = sum(state_arr2 - predict_loess$fit)
  lower95 = sum(state_arr2 - df_plots$lower[(length(df_plots$lower) - 17):length(df_plots$lower)])
  upper95 = sum(state_arr2 - df_plots$upper[(length(df_plots$upper) -17):length(df_plots$upper)])

  excess2 = sum(state_arr2[1:4] - predict_loess$fit[1:4])
  lower952 = sum(state_arr2[1:4] - df_plots$lower[(length(df_plots$lower) - 17):(length(df_plots$lower)-14)])
  upper952 = sum(state_arr2[1:4] - df_plots$upper[(length(df_plots$upper) -17):(length(df_plots$upper)-14)])
  
  df_excess_loess[i, ] = c(state_name, excess, lower95, upper95, excess2, lower952, upper952)
  
  g <- ggplot(df_plots, aes(x=x,y=mean, ymin=lower, ymax=upper, colour= "LOESS model")) + 
  geom_ribbon(color = NA,#border line color
                fill="#c1ef7d") + 
  geom_line(colour="#94cb00") + 
  geom_vline(xintercept = 2020.12, linetype = "dashed", color = "black", size = 0.8)+
  geom_line(aes(y=actual, color = "Actual data")) + 
  xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) +
  theme_bw() + theme(legend.position="none", legend.title = element_blank()) +
  scale_color_manual(values = c("blueviolet", "chartreuse3")) + 
  guides(color=guide_legend("Time Period")) + scale_x_continuous(breaks=c(2015, 2016, 2017, 2018, 2019, 2020, 2021))

 plots_loess[[i]] = g
}

```

```{r}
df_excess_loess = merge(df_excess_loess, state_pop)
df_excess_loess = df_excess_loess %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000), Excess_pop_2 = as.numeric(Excess_2)/(Population/100000), Lower95_pop_2 =as.numeric(Lower95_2)/(Population/100000), Upper95_pop_2 =as.numeric(Upper95_2)/(Population/100000))
colnames(df_excess_loess)[1] = 'state'
```

```{r}
# Save Plots
for (i in seq(1, length(plots_loess))){
  ggsave(paste('plots/states_loess/',state_names[i]),
  plot = plots_loess[[i]],
  device = "png", width = 6, height = 5, dpi = 600)
}
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_loess %>% mutate(state = tolower(state))
df_excess_graph = df_excess_graph %>% filter(state != 'district of columbia')
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Loess"),
  device = "png")
```

```{r}
# Plot of values and error bars for loess
df_excess_loess = df_excess_loess[order(df_excess_loess$Excess_pop, decreasing = T),]
df_excess_loess$state = factor(df_excess_loess$state, levels=df_excess_loess[order(df_excess_loess$Excess_pop), "state"])
vert_plot <- ggplot(
    df_excess_loess,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Loess"),
  device = "png", width = 7, height = 9)
```

# Comparison between Estimates

```{r}
df_c = df_excess %>% select(state, Excess_pop) %>% rename(e1 = Excess_pop)
df_l = df_excess_loess %>% select(state, Excess_pop) %>% rename(e2 = Excess_pop)
df_compare = merge(df_c, df_l)
df_compare$diff = df_compare$e1 - df_compare$e2
g <- ggplot(df_compare) +
  geom_boxplot(aes(y = diff)) +
  scale_x_discrete( ) +
  labs(y = "sARIMA Estimate minus LOESS Estimate") +
  theme_clean() + 
  coord_fixed(ratio = 0.05)
g
ggsave(paste('plots/',"comparison"),
  device = "png")
```

```{r}
wilcox.test(df_compare$e1, df_compare$e2, paired = TRUE)
```


# County Level Estimates

```{r}
# Preprocessing
county = county_1[county_1$Notes == "", ]
# Not Including 2021 for now
# county_2 = county_2[county_2$Notes == "" & county_2$Year.Code == 2021, ]
# colnames(county_2)[2] <- 'County'
# colnames(county_2)[3] <- 'County.Code'
# county = rbind(county_1, county_2)

'%!in%' <- function(x,y)!('%in%'(x,y))
#Remove 2010, 2011 data to get more counties
# county = county[(county$Year != '2010')&(county$Year != '2011'),]
ridCode = unique(county$County.Code[(county$Deaths == 'Suppressed')|(county$Deaths == 'Missing')])
county = county[county$County.Code %!in% ridCode, ]
county$Deaths = as.numeric(levels(county$Deaths))[county$Deaths]
```

## ARIMA
```{r}
county_codes = unique(county$County.Code)
df_excess_county <- data.frame(matrix(nrow = length(county_codes), ncol = 5))
colnames(df_excess_county) <- c('County', 'Code', 'Excess', 'Lower95', 'Upper95')
j = 1
for (i in county_codes){
  county_data = county %>% filter(County.Code == i)
  t_county = ts(county_data$Deaths[1:(length(county_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  # Replaced by full parameter search
  # fit <- tryCatch({
  #       Arima(t_county, order = c(1, 1, 1), method="ML")
  #   },
  #   error = function(e){
  #     Arima(t_county, order = c(2, 2, 2), method="ML")
  #   })
  f_fit<-forecast(fit, robust = TRUE, h = 1, level = 95)
  true_2020 = county_data$Deaths[length(county_data$Deaths)]
    g<-autoplot(t_county, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(as.character(unique(county_data$County))) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom") + 
    geom_point(aes(x=2020, y=true_2020), colour="blue")
    ggsave(paste('plots/counties/',as.character(unique(county_data$County))),
  plot = g,
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_county[j, ] = c(as.character(unique(county_data$County)), i, excess, lower95, upper95)
  j = j + 1
}
```

```{r}
# Actually looping through all parameters
county_codes = unique(county$County.Code)
df_excess_county <- data.frame(matrix(nrow = length(county_codes), ncol = 5))
colnames(df_excess_county) <- c('County', 'Code', 'Excess', 'Lower95', 'Upper95')
k = 1
parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 2))
for (i in county_codes){
  county_data = county %>% filter(County.Code == i)
  t_county = ts(county_data$Deaths[1:(length(county_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_county, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_county, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(unique(county_data$County))
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  
  true_2020 = county_data$Deaths[length(county_data$Deaths)]
    g<-autoplot(t_county, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(as.character(unique(county_data$County))) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom") + 
    geom_point(aes(x=2020, y=true_2020), colour="blue")
    ggsave(paste('plots/counties/',as.character(unique(county_data$County))),
  plot = g,
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_county[k, ] = c(as.character(unique(county_data$County)), i, excess, lower95, upper95)
  k = k + 1
}
```


```{r}
# Edit and Merge on FIPS
for (row in 1:nrow(df_excess_county)){
  if (nchar(df_excess_county[row, 'Code']) == 4){
    newChar = paste0("0", df_excess_county[row, 'Code'])
    df_excess_county[row, 'Code'] = newChar
  }
}
df_excess_county <- merge(df_excess_county, county_pop, by.x = 'Code', by.y = 'fips')
```

```{r}
# Pop Normalization
df_excess_county = df_excess_county %>% mutate(Excess_pop = as.numeric(Excess)/(Pop/100000), Lower95_pop = as.numeric(Lower95)/(Pop/100000), Upper95_pop = as.numeric(Upper95)/(Pop/100000))
colnames(df_excess_county)[1] = 'fips'
colnames(df_excess_county)[2] = 'County'
```


```{r}
# County Level US plot
plot_usmap(data = df_excess_county, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma, na.value = "white") + 
  labs(title = "County-Level Excess Deaths Estimates") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map Counties"),
  device = "png")
```

```{r}
load('data/counties.rda')
load('data/states.rda')
df_excess_graphc <- df_excess_county
countyGraph = merge (counties, df_excess_graphc, by.x = 'county_fips', by.y = 'fips', all.x = TRUE)
countyGraph = countyGraph[order(countyGraph$county_fips, countyGraph$order), ]
countyGraph = countyGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "gray50"),
    size = ifelse(outline == "gray50", 0.06, 0.25)
  )

ggplot(countyGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 geom_polygon(data = states, mapping = aes(long, lat, group = group),
               fill = NA, color = "gray30", size = 0.15) +
 scale_fill_gradient2(
    low = "purple2",
    mid = "white",
    high = "#ff7800",
    midpoint = 0,
    na.value = "grey85", name = "Excess Deaths per 100,000"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map Counties"),
  device = "png", dpi =1000)
```

```{r}
df_excess_county = df_excess_county[order(df_excess_county$Excess_pop, decreasing = T),]
df_excess_county$County = factor(df_excess_county$County, levels=df_excess_county[order(df_excess_county$Excess_pop), "County"])
vert_plot <- ggplot(
    df_excess_county[1:50,],
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = County)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("County Excess Deaths per 100,000 Persons")+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot County"),
  device = "png", width = 7, height = 9)
```

## LOESS test (Not used in final version)
```{r}
county_data = county %>% filter(County.Code == "54099")
loess_model <- loess(county_data$Deaths[1:(length(county_data$Deaths) -1)] ~ as.numeric(1:(length(county_data$Deaths)-1)), control = loess.control(surface = "direct"))
fit_loess = predict(loess_model,se=TRUE)
predict_loess = predict(loess_model,as.numeric((length(county_data$Deaths) + 1):(length(county_data$Deaths) + 1)), se = TRUE)
x = linspace(2010, 2020, length(county_data$Deaths))

df_plots <- data.frame(matrix(nrow = length(county_data$Deaths), ncol = 4))
colnames(df_plots) <- c('x', 'mean', 'upper', 'lower')

df_plots$x <- x
df_plots$mean = c(fit_loess$fit, predict_loess$fit)
df_plots$lower = c((fit_loess$fit - 1.96*fit_loess$se.fit), (predict_loess$fit - 1.96*predict_loess$se.fit))
df_plots$upper = c((fit_loess$fit + 1.96*fit_loess$se.fit), (predict_loess$fit + 1.96*predict_loess$se.fit))
df_plots$actual = county_data$Deaths

ggplot(df_plots, aes(x=x,y=mean, ymin=lower, ymax=upper)) + 
geom_line() + 
geom_ribbon(alpha=0.3) + 
geom_point(aes(y=actual)) + 
theme_bw() + ggtitle(as.character(unique(county_data$County))) + ylab('Deaths') + xlab('Years') +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
```

# Regression Analysis
## CCVI
```{r}
ccvi_state <- read_excel('data/surgo_ccvi/ccvi.xlsx')
ccvi_state$stateName <- str_to_title(ccvi_state$stateName)
ccvi_state$stateName[9] = "District of Columbia"
```

```{r}
df_excess_predictors <- merge(df_excess, ccvi_state, by.x = 'state', by.y = 'stateName')
df_excess_predictors <- merge(df_excess_predictors, state_pop, by.x = 'state', by.y = 'State')
```


```{r}
# Try both Excess_pop 2 and Excess_pop
wls_model_ccvi <- lm(Excess_pop_2 ~ theme1 + theme2 + theme3 + theme5 + theme6 +theme7, data = df_excess_predictors, weights = log(Population.x))
summary(wls_model_ccvi)
```

### County CCVI

```{r}
ccvi_indicators <- read_excel('data/surgo_ccvi/ccvi_county_raw_indicators_public.xlsx')
ccvi_county <- read_excel('data/surgo_ccvi/ccvi.xlsx', sheet = 'ccvi-US-county')
ccvi_county$FIPS = as.character(ccvi_county$FIPS)
ccvi_indicators$FIPS = as.character(ccvi_indicators$FIPS)
```

```{r}
for (row in 1:nrow(ccvi_county)){
  if (nchar(ccvi_county[row, 'FIPS']) == 4){
    newChar = paste0("0", ccvi_county[row, 'FIPS'])
    ccvi_county[row, 'FIPS'] = newChar
  }
}
```

```{r}
for (row in 1:nrow(ccvi_indicators)){
  if (nchar(ccvi_indicators[row, 'FIPS']) == 4){
    newChar = paste0("0", ccvi_indicators[row, 'FIPS'])
    ccvi_indicators[row, 'FIPS'] = newChar
  }
}
```

```{r}
df_excess_predictors_county <- merge(df_excess_county, ccvi_county, by.x = 'fips', by.y = 'FIPS')
df_excess_predictors_county <- merge(df_excess_predictors_county, ccvi_indicators, by.x = 'fips', by.y = 'FIPS')
```

```{r}
wls_model_ccvi_county <- lm(Excess_pop ~ theme1 + theme2 + theme3 + theme5 + theme6 +theme7, data = df_excess_predictors_county, weights = log(Pop))
summary(wls_model_ccvi_county)
```

```{r}
wls_model_ccvi_county <- lm(Excess_pop ~ ccvi, data = df_excess_predictors_county, weights = log(Pop))
summary(wls_model_ccvi_county)
```

```{r}
wls_model_ccvi_county <- lm(Excess_pop ~ ses_EP_POV + ses_EP_UNEMP + ses_EP_PCI + ses_EP_NOHSDP + ses_EP_UNINSUR + ms_EP_MINRTY + hh_EP_CROWD + pm_EP_AGE65 + hc_icu_capita + hc_hosp_beds_capita + hc_PHEP_capita + hc_spend_capita + hc_epi_capita + hc_health_labs_capita + hc_emer_services_capita + hc_pqi + hh_EP_SNGPNT + hh_EP_AGE17, data = df_excess_predictors_county, weights = log(Pop))
summary(wls_model_ccvi_county)
```

## SVI (Results not Included)

```{r}
df_svi <- read.csv('data/SVI2018_US_COUNTY.csv')
for (row in 1:nrow(df_svi)){
  if (nchar(df_svi[row, 'FIPS']) == 4){
    newChar = paste0("0", df_svi[row, 'FIPS'])
    df_svi[row, 'FIPS'] = newChar
  }
}
```

```{r}
df_excess_predictors_county <- merge(df_excess_county, df_svi, by.x = 'fips', by.y = 'FIPS')
```

```{r}
wls_model_svi_county <- lm(Excess_pop ~ RPL_THEME1 + RPL_THEME2 + RPL_THEME3 + RPL_THEME4, data = df_excess_predictors_county, weights = log(Pop))
summary(wls_model_svi_county)
```

# Basic Visualizations based on Demographics

```{r}
# Import dataset shared by Prof Therriault
df_dem <- read.csv('data/dem_data.csv')
df_dem$County = as.character(levels(df_dem$County))[df_dem$County]
df_dem$State = as.character(levels(df_dem$State))[df_dem$State]
for (row in 1:nrow(df_dem)) {
  if (row %% 100 == 0){
    print(row)
  }
    cname <- df_dem[row, "County"]
    sname = df_dem[row, "State"]
    tryfips <- tryCatch({
        fips(state = substring(sname, 2, nchar(sname)), county = cname)
    },
    error = function(e){
      NA
    })
    df_dem[row, 'fips'] <- tryfips
}
```


```{r}
df_excess_dem = merge(df_excess_county, df_dem, by = 'fips')
df_excess_dem <- subset (df_excess_dem , select = -County.y)
colnames(df_excess_dem)[2] <- 'county'
```

```{r}
df_excess_dem = df_excess_dem[order(df_excess_dem$black_pct, decreasing = T),]
b <- df_excess_dem$Excess_pop[1:20]
df_excess_dem = df_excess_dem[order(df_excess_dem$white_pct, decreasing = T),]
w <- df_excess_dem$Excess_pop[1:20]
df_excess_dem = df_excess_dem[order(df_excess_dem$aian_pct, decreasing = T),]
aian <- df_excess_dem$Excess_pop[1:20]
df_excess_dem = df_excess_dem[order(df_excess_dem$nhpi_pct, decreasing = T),]
nh <- df_excess_dem$Excess_pop[1:20]
```


```{r}
df_top_race <- data.frame(
  values = c(b, w, aian, nh),
  Race = rep(c("Black", "White", "American Indian and Alaskan Native", "Native Hawaiin and Pacific Islander"), each = 20)
  )
```

```{r}
ggplot(df_top_race, aes(x = values, fill = Race)) + geom_density(alpha = 0.5) + theme_clean()+
  xlab('Drug-Related Excess Mortality per 100,000 Persons')
```



# American Community Project
```{r}
# Key	New Names
# 1	Exurbs
# 2	Graying America
# 3	African American South
# 4	Evangelical Hubs
# 5	Working Class Country
# 6	Military Posts
# 7	Urban Suburbs
# 8	Hispanic Centers
# 9	Native American Lands
# 10	Rural Middle America
# 11	College Towns
# 12	LDS Enclaves
# 13	Aging Farmlands
# 14	Big Cities
# 15	Middle Suburbs
acs_names = c('Exurbs', 'Graying America', 'African American South', 'Evangelical Hubs', 'Working Class Country', 'Military Posts', 'Urban Suburbs', 'Hispanic Centers', 'Native American Lands', 'Rural Middle America', 'College Towns', 'LDS Enclaves', 'Aging Farmlands', 'Big Cities', 'Middle Suburbs')
```

```{r}
# Calculate Mean and SD for each County Type
acs <- read_excel('data/County-Type.xlsx')
acs$fips = as.character(acs$fips)
for (row in 1:nrow(acs)){
  if (nchar(acs[row, 'fips']) == 4){
    newChar = paste0("0", acs[row, 'fips'])
    acs[row, 'fips'] = newChar
  }
}
df_excess_acs = merge(df_excess_county, acs, by = 'fips')
df_excess_acs <- subset (df_excess_acs, select = -County.y)

df_mean_acs = df_excess_acs %>%  group_by(Type_Number) %>% summarise(mean = mean(Excess_pop), sd = sd(Excess_pop))
df_mean_acs$Type_Number = as.factor(df_mean_acs$Type_Number)
```


```{r}
# Plot results
p<-ggplot(data=df_mean_acs, aes(x=Type_Number, y=mean)) +
  geom_bar(stat="identity", fill="steelblue") + 
  xlab('County Type') +
  ylab('Mean Excess Deaths per 100,000')+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1,
                 position=position_dodge(.9)) + 
  scale_x_discrete(breaks=c("1","2","3", '4', '5', '6', '7', '8', '9', '10', '11', '12', '13', '14', '15'), labels=acs_names) + 
  coord_flip() + 
  theme_clean()
p
ggsave(paste('plots/',"ACP Excess"),
  device = "png")
```

# Repeat of Original Analysis for Specific Drug Types

## Opioids
```{r}
state_1 <- read.delim("data/All Opioids/State_Month.txt")
state_2 <- read.delim("data/All Opioids/State_Month_2.txt")
state_1 = state_1[state_1$Notes == "", ]
state_2 = state_2[state_2$Notes == "" & state_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
colnames(state_2)[2] <- 'State'
colnames(state_2)[3] <- 'State.Code'
state = rbind(state_1, state_2)
```

```{r}
plots = list()
state_names = as.character(unique(state$State))
df_params_op <- data.frame(matrix(nrow = 51, ncol = 7))
colnames(df_params_op) <- c('State', 'p', 'd', 'q', 'P', 'D', 'Q')
df_excess_op <- data.frame(matrix(nrow = 51, ncol = 7))
colnames(df_excess_op) <- c('State', 'Excess', 'Lower95', 'Upper95', 'Excess_2', 'Lower95_2', 'Upper95_2')
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_State = ts(State_data$Deaths[1:(length(State_data$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)
  
  parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 2),
              P = seq(0, 2), D = seq(0, 2), Q = seq(0, 2))
  
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    
    fit <- tryCatch(
    {
        Arima(t_State, order = c(parameters$p[j], parameters$d[j], parameters$q[j]),
                      seasonal = c(parameters$P[j], parameters$D[j], parameters$Q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_State, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(state_name)
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  if (break_tag){
      next
    }

  f_fit<-forecast(best_fit, robust = TRUE, h = 18, level = 95)
  
  t_State2 = ts(State_data$Deaths[(length(State_data$Deaths) - 17):length(State_data$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)
  excess = as.numeric(t_State2) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(t_State2) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(t_State2) - as.numeric(f_fit[["upper"]])
  
  excess2 = as.numeric(t_State2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]
  lower952 = as.numeric(t_State2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]
  upper952 = as.numeric(t_State2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]

  df_params_op[i, ] = c(state_name, best_params$p, best_params$d, best_params$q, best_params$P, best_params$D, best_params$Q)
  df_excess_op[i, ] = c(state_name, sum(excess), sum(lower95), sum(upper95), sum(excess2), sum(lower952), sum(upper952))
  
model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
 g = autoplot(t_State, series="Data") + 
     autolayer(fit$fitted, series=model_name) +
     autolayer(f_fit, series="Prediction") +
     autolayer(t_State2, series = "True Post-Pandemic") +
     xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
 plots[[i]] = g
}
```

```{r}
df_excess_op = merge(df_excess_op, state_pop)
df_excess_op = df_excess_op %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000), Excess_pop_2 = as.numeric(Excess_2)/(Population/100000), Lower95_pop_2 =as.numeric(Lower95_2)/(Population/100000), Upper95_pop_2 =as.numeric(Upper95_2)/(Population/100000))
colnames(df_excess_op)[1] = 'state'
```

```{r}
plot_usmap(data = df_excess_op, values = "Excess_pop", color = "black") + 
  scale_fill_gradient2(
    low = "blue", 
    mid = "white", 
    high = "red", 
    midpoint = 0
  ) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic- Opioids") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States Opioids"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

## Synthetic Opioids
```{r}
state <- read.delim("data/Synthetic Opioids/State_Year.txt")
plots = list()
df_excess_synth <- data.frame(matrix(nrow = 51, ncol = 4))
k = 1
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_state = ts(State_data$Deaths[1:(length(State_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_state, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_state, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(unique(county_data$County))
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  true_2020 = State_data$Deaths[length(State_data$Deaths)]
    g<-autoplot(t_state, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
        geom_point(aes(x=2020, y=true_2020), colour="blue") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
  ggsave(paste('plots/states_synth/',state_names[i]),
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_synth[k, ] = c(state_name, excess, lower95, upper95)
  k = k + 1
}
```

```{r}
colnames(df_excess_synth) <- c('State', 'Excess', 'Lower95', 'Upper95')
df_excess_synth = merge(df_excess_synth, state_pop)
df_excess_synth = df_excess_synth %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000))
colnames(df_excess_synth)[1] = 'state'
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_synth %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(fill = Excess_pop, color = outline, size = size)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Synthetic Opioids"),
  device = "png")
```


```{r}
df_excess_synth = df_excess_synth[order(df_excess_synth$Excess_pop, decreasing = T),]
df_excess_synth$state = factor(df_excess_synth$state, levels=df_excess_synth[order(df_excess_synth$Excess_pop), "state"])
df_excess_synth = df_excess_synth[complete.cases(df_excess_synth), ]
vert_plot <- ggplot(
    df_excess_synth,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  xlab('Excess Deaths per 100,000 Persons')+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Synthetic Opioids"),
  device = "png", width = 7, height = 9)
```

```{r}
country_1 <- read.delim("data/Synthetic Opioids/Country_Month.txt")
country_2 <- read.delim("data/Synthetic Opioids/Country_Month_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

```{r}
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- tryCatch(
    {
        Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                      seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
    },
    error = function(e){
      
        Arima(t_Country, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      })
  
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
autoplot(t_Country, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Post-Pandemic") +
  xlab("Year") + ylab("Synthetic Opioids Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
ggsave(paste('plots/',"US States Time Synthetic Opioids"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

## Cocaine
```{r}
state <- read.delim("data/Cocaine/State_Year.txt")
plots = list()
df_excess_coc <- data.frame(matrix(nrow = 51, ncol = 4))
k = 1
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_state = ts(State_data$Deaths[1:(length(State_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  if (mean(is.na(as.numeric(t_state))) > 0.7){
    next
  }
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_state, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_state, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(state_name)
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  true_2020 = State_data$Deaths[length(State_data$Deaths)]
    g<-autoplot(t_state, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
        geom_point(aes(x=2020, y=true_2020), colour="blue") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
  ggsave(paste('plots/states_cocaine/',state_names[i]),
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_coc[k, ] = c(state_name, excess, lower95, upper95)
  k = k + 1
}
```

```{r}
colnames(df_excess_coc) <- c('State', 'Excess', 'Lower95', 'Upper95')
df_excess_coc = merge(df_excess_coc, state_pop)
df_excess_coc = df_excess_coc %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000))
colnames(df_excess_coc)[1] = 'state'
```

```{r}
plot_usmap(data = df_excess_coc, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic- Cocaine") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States Cocaine"),
  device = "png")
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_coc %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Cocaine"),
  device = "png")
```

```{r}
df_excess_coc = df_excess_coc[order(df_excess_coc$Excess_pop, decreasing = T),]
df_excess_coc$state = factor(df_excess_coc$state, levels=df_excess_coc[order(df_excess_coc$Excess_pop), "state"])
vert_plot <- ggplot(
    df_excess_coc,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  xlab('Excess Deaths per 100,000 Persons')+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Cocaine"),
  device = "png", width = 7, height = 9)
```


```{r}
country_1 <- read.delim("data/Cocaine/Country_Year.txt")
country_2 <- read.delim("data/Cocaine/Country_Year_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

```{r}
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                    seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
autoplot(t_Country, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Post-Pandemic") +
  xlab("Year") + ylab("Cocaine Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
ggsave(paste('plots/',"US States Time Cocaine"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```


## Methamphetamines
```{r}
state <- read.delim("data/Meth/State_Year.txt")
df_excess_meth <- data.frame(matrix(nrow = 51, ncol = 4))
k = 1
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_state = ts(State_data$Deaths[1:(length(State_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  if (mean(is.na(as.numeric(t_state))) > 0.7){
    next
  }
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_state, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_state, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(state_name)
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  true_2020 = State_data$Deaths[length(State_data$Deaths)]
    g<-autoplot(t_state, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
        geom_point(aes(x=2020, y=true_2020), colour="blue") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
  ggsave(paste('plots/states_meth/',state_names[i]),
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_meth[k, ] = c(state_name, excess, lower95, upper95)
  k = k + 1
}
```

```{r}
colnames(df_excess_meth) <- c('State', 'Excess', 'Lower95', 'Upper95')
df_excess_meth = merge(df_excess_meth, state_pop)
df_excess_meth = df_excess_meth %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000))
colnames(df_excess_meth)[1] = 'state'
```

```{r}
plot_usmap(data = df_excess_meth, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic- Methamphetamines") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States Meth"),
  device = "png")
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_meth %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Meth"),
  device = "png")
```


```{r}
df_excess_meth = df_excess_meth[order(df_excess_meth$Excess_pop, decreasing = T),]
df_excess_meth$state = factor(df_excess_meth$state, levels=df_excess_meth[order(df_excess_meth$Excess_pop), "state"])

vert_plot <- ggplot(
    df_excess_meth,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  xlab('Excess Deaths per 100,000 Persons')+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Meth"),
  device = "png", width = 7, height = 9)
```

```{r}
country_1 <- read.delim("data/Meth/Country_Month.txt")
country_2 <- read.delim("data/Meth/Country_Month_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

```{r}
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                    seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
autoplot(t_Country, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Post-Pandemic") +
  xlab("Year") + ylab("Methamphetamines Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
ggsave(paste('plots/',"US States Time Methamphetamines"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

## Benzadiapenes
```{r}
state <- read.delim("data/Benz/State_Year.txt")
plots = list()
df_excess_benz <- data.frame(matrix(nrow = 51, ncol = 4))
k = 1
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_state = ts(State_data$Deaths[1:(length(State_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  if (mean(is.na(as.numeric(t_state))) > 0.7){
    next
  }
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_state, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_state, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(state_name)
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  true_2020 = State_data$Deaths[length(State_data$Deaths)]
    g<-autoplot(t_state, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
        geom_point(aes(x=2020, y=true_2020), colour="blue") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
  ggsave(paste('plots/states_benz/',state_names[i]),
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_benz[k, ] = c(state_name, excess, lower95, upper95)
  k = k + 1
}
```

```{r}
colnames(df_excess_benz) <- c('State', 'Excess', 'Lower95', 'Upper95')
df_excess_benz = merge(df_excess_benz, state_pop)
df_excess_benz= df_excess_benz %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000))
colnames(df_excess_benz)[1] = 'state'
```

```{r}
plot_usmap(data = df_excess_benz, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic- Benzodiapenes") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States Benzodiazepenes"),
  device = "png")
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_benz %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Benzodiazepenes"),
  device = "png")
```


```{r}
df_excess_benz = df_excess_benz[order(df_excess_benz$Excess_pop, decreasing = T),]
df_excess_benz$state = factor(df_excess_benz$state, levels=df_excess_benz[order(df_excess_benz$Excess_pop), "state"])
df_excess_benz = df_excess_benz[complete.cases(df_excess_benz), ]
vert_plot <- ggplot(
    df_excess_benz,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  xlab('Excess Deaths per 100,000 Persons')+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Benz"),
  device = "png", width = 7, height = 9)
```

```{r}
country_1 <- read.delim("data/Benz/Country_Month.txt")
country_2 <- read.delim("data/Benz/Country_Month_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

```{r}
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                    seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
autoplot(t_Country, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Post-Pandemic") +
  xlab("Year") + ylab("benzodiazepines Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
ggsave(paste('plots/',"US States Time benzodiazepines"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

## Alcohol

```{r}
state <- read.delim("data/Alcohol/State_Year.txt")
df_excess_alc <- data.frame(matrix(nrow = 51, ncol = 4))
k = 1
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(State_data$Deaths)
  t_state = ts(State_data$Deaths[1:(length(State_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  if (mean(is.na(as.numeric(t_state))) > 0.7){
    next
  }
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_state, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_state, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(state_name)
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  true_2020 = State_data$Deaths[length(State_data$Deaths)]
    g<-autoplot(t_state, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
        geom_point(aes(x=2020, y=true_2020), colour="blue") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
  ggsave(paste('plots/states_alc/',state_names[i]),
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_alc[k, ] = c(state_name, excess, lower95, upper95)
  k = k + 1
}
```

```{r}
colnames(df_excess_alc) <- c('State', 'Excess', 'Lower95', 'Upper95')
df_excess_alc = merge(df_excess_alc, state_pop)
df_excess_alc= df_excess_alc %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000))
colnames(df_excess_alc)[1] = 'state'
```

```{r}
plot_usmap(data = df_excess_alc, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic- Alcohol") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States Alcohol"),
  device = "png")
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_alc %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Alcohol"),
  device = "png")
```

```{r}
df_excess_alc = df_excess_alc[order(df_excess_alc$Excess_pop, decreasing = T),]
df_excess_alc$state = factor(df_excess_alc$state, levels=df_excess_alc[order(df_excess_alc$Excess_pop), "state"])
vert_plot <- ggplot(
    df_excess_alc,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  xlab('Excess Deaths per 100,000 Persons')+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Alcohol"),
  device = "png", width = 7, height = 9)
```

```{r}
country_1 <- read.delim("data/Alcohol/Country_Month.txt")
country_2 <- read.delim("data/Alcohol/Country_Month_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

```{r}
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                    seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
autoplot(t_Country, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Post-Pandemic") +
  xlab("Year") + ylab("Alcohol Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
ggsave(paste('plots/',"US States Time Alcohol"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

## Heroin

```{r}
state <- read.delim("data/Heroin/State_Year.txt")
plots = list()
df_excess_her <- data.frame(matrix(nrow = 51, ncol = 4))
k = 1
for (i in seq(1, length(state_names))){
  state_name = state_names[i]
  State_data = state %>% filter(State == state_name)
  State_data[State_data == 'Suppressed'] <- NA
  State_data$Deaths = as.numeric(levels(State_data$Deaths))[State_data$Deaths]
  t_state = ts(State_data$Deaths[1:(length(State_data$Deaths) -1)], start = decimal_date(as.Date("2010-01-01")))
  if (mean(is.na(as.numeric(t_state))) > 0.7){
    next
  }
  best_aic = 10000
  break_tag = FALSE
  for (j in seq(1, nrow(parameters))){
    paste(parameters$p[j], parameters$d[j], parameters$q[j])
    fit <- tryCatch(
    {
        Arima(t_state, order = c(parameters$p[j], parameters$d[j], parameters$q[j]), method="ML")
    },
    error = function(e){
      
      tryCatch(
      {
        Arima(t_state, order = c(1, 1, 0),
                      seasonal = c(1, 0, 1), method="ML")
      },
      error = function(e){
        TRUE
        break_tag <<- TRUE
      }
      )
      
    }
    )
    if (break_tag){
      print("hi")
      print(state_name)
      next
    }
    
    aic = fit[["aicc"]]
    if (aic<best_aic){
      best_aic = aic
      best_fit = fit
      best_params = parameters[j, ]
    }
  }
  # if (break_tag){
  #     next
  #   }

  f_fit<-forecast(best_fit, robust = TRUE, h = 1, level = 95)
  true_2020 = State_data$Deaths[length(State_data$Deaths)]
    g<-autoplot(t_state, series="Data") + 
         autolayer(fit$fitted, series="SARIMA") +
         autolayer(f_fit, series="Prediction") +
        geom_point(aes(x=2020, y=true_2020), colour="blue") +
         xlab("Year") + ylab("Drug Overdose Deaths") + ggtitle(state_name) + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
  ggsave(paste0('plots/states_heroin/',state_names[i]),
  device = "png")
  excess = as.numeric(true_2020) - as.numeric(f_fit[["mean"]])
  lower95 = as.numeric(true_2020) - as.numeric(f_fit[["lower"]])
  upper95 = as.numeric(true_2020) - as.numeric(f_fit[["upper"]])
  df_excess_her[k, ] = c(state_name, excess, lower95, upper95)
  k = k + 1
}
```

```{r}
colnames(df_excess_her) <- c('State', 'Excess', 'Lower95', 'Upper95')
df_excess_her = merge(df_excess_her, state_pop)
df_excess_her= df_excess_her %>% mutate(Excess_pop = as.numeric(Excess)/(Population/100000), Lower95_pop =as.numeric(Lower95)/(Population/100000), Upper95_pop =as.numeric(Upper95)/(Population/100000))
colnames(df_excess_her)[1] = 'state'
```

```{r}
plot_usmap(data = df_excess_her, values = "Excess_pop", color = "black") + 
  scale_fill_viridis_c(option= "magma", name = "Excess Deaths Per 100,000 Persons", label = scales::comma) + 
  labs(title = "US Excess Drug Overdose Deaths During the Pandemic- Heroin") +
  theme(legend.position = "right")
ggsave(paste('plots/',"US Map States Heroin"),
  device = "png")
```

```{r}
library(fiftystater)
statesGraph <- fifty_states
df_excess_graph = df_excess_her %>% mutate(state = tolower(state))
statesGraph = merge (statesGraph, df_excess_graph, by.x = 'id', by.y = 'state', all.x = TRUE)
statesGraph = statesGraph[order(statesGraph$id, statesGraph$order), ]
statesGraph = statesGraph %>% mutate(
    outline = case_when(
      Upper95_pop > 0 ~ "#ff5500",
      TRUE ~ "black"),
    size = ifelse(outline == "black", 0.25, 1)
  )

ggplot(statesGraph, aes(long, lat, group = group)) +
 geom_polygon(aes(color = outline, size = size, fill = Excess_pop)) +
 scale_fill_gradient2(
    low = "purple2", 
    mid = "white", 
    high = "#ff7800", 
    midpoint = 0, 
    na.value = "grey80", name = "Excess Deaths per 100k"
  ) +
 scale_color_identity() +
 scale_size_identity() +
 theme_void() +
theme(legend.position = "bottom")
ggsave(paste('plots/',"US Map States Heroin"),
  device = "png")
```

```{r}
df_excess_her = df_excess_her[order(df_excess_her$Excess_pop, decreasing = T),]
df_excess_her$state = factor(df_excess_her$state, levels=df_excess_her[order(df_excess_her$Excess_pop), "state"])
df_excess_her = df_excess_her[complete.cases(df_excess_her), ]
vert_plot <- ggplot(
    df_excess_her,
    aes(x = Excess_pop,xmin = Upper95_pop,xmax = Lower95_pop, y = state)) +
    geom_errorbarh(height = .2, 
                   position = position_dodge(width = .7),
                   alpha = .9) +
    geom_point(position = position_dodge(width = .7),
               alpha = .9) +
    geom_vline(xintercept = 0, 
               color = "black",
               alpha = .7) + 
    ggtitle("State Excess Deaths per 100,000 Persons")+
  xlab('Excess Deaths per 100,000 Persons')+
  theme_clean()
vert_plot
ggsave(paste('plots/',"Vertical Plot Heroin"),
  device = "png", width = 7, height = 9)
```

```{r}
country_1 <- read.delim("data/Heroin/Country_Month.txt")
country_2 <- read.delim("data/Heroin/Country_Month_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
```

```{r}
country[country == 'Suppressed'] <- NA
country$Deaths = as.numeric(country$Deaths)

t_Country = ts(country$Deaths[1:(length(country$Deaths) - 18)], start = decimal_date(as.Date("2015-01-01")), frequency = 12)

parameters = expand.grid(p = seq(0, 2), d = seq(0, 2), q = seq(0, 1),
            P = seq(0, 1), D = seq(0, 1), Q = seq(0, 1))

best_aic = 10000
for (i in seq(1, nrow(parameters))){
  fit <- Arima(t_Country, order = c(parameters$p[i], parameters$d[i], parameters$q[i]),
                    seasonal = c(parameters$P[i], parameters$D[i], parameters$Q[i]), method="ML")
  aic = fit[["aicc"]]
  if (aic<best_aic){
    best_aic = aic
    best_fit = fit
    best_params = parameters[i, ]
  }
  
}
# fit <- Arima(t_Country, order = c(1, 1, 1),
#                      seasonal = c(1, 0, 1), method="ML")
f_fit<-forecast(best_fit, h = 18, robust = TRUE, level = 95)

# Confidence Interval Calculation Attempt 1
# sim <- matrix(NA, ncol=18,nrow=1000)
# for(i in 1:1000)
#   sim[i,] <- simulate(fit,18, future=TRUE)
# se <- apply(sim,2,sd)/sqrt(1000)
# f_fit$upper <- f_fit$mean + 1.96*se
# f_fit$lower <- f_fit$mean - 1.96*se

# Confidence Interval Attempt 2
# f_fit$upper <- f_fit$mean + 1.96*sqrt(fit$sigma2)
# f_fit$lower <- f_fit$mean - 1.96*sqrt(fit$sigma2)

t_Country2 = ts(country$Deaths[(length(country$Deaths) - 17):length(country$Deaths)], start = decimal_date(as.Date("2020-03-01")), frequency = 12)

model_name = paste('SARIMA(', best_params[1], best_params[2], best_params[3], ')(', best_params[4], best_params[5], best_params[6], ')[12]')
autoplot(t_Country, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(t_Country2, series = "True Post-Pandemic") +
  xlab("Year") + ylab("Heroin Overdose Deaths") + ggtitle("United States") + theme_bw()+theme(legend.title = element_blank(),legend.position = "bottom")
ggsave(paste('plots/',"US States Time Alcohol"),
  device = "png")
```

```{r}
US_pop_norm = (state_pop$Population[state_pop$State == 'United States']/100000)
excess = sum(as.numeric(t_Country2) - as.numeric(f_fit[["mean"]])) / US_pop_norm
lower95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["lower"]])) / US_pop_norm
upper95 = sum(as.numeric(t_Country2) - as.numeric(f_fit[["upper"]])) / US_pop_norm

excess2 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["mean"]])[1:4]) / US_pop_norm
lower952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["lower"]])[1:4]) / US_pop_norm
upper952 = sum(as.numeric(t_Country2)[1:4] - as.numeric(f_fit[["upper"]])[1:4]) / US_pop_norm
```

# Create Tables

# State Table
```{r}
excess_small <- df_excess %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename(Overall = Excess_pop)
alc_small <- df_excess_alc %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename(Alcohol = Excess_pop)
benz_small <- df_excess_benz %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename(Benzodiapenes = Excess_pop)
coc_small <- df_excess_coc %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename(Cocaine = Excess_pop)
her_small <- df_excess_her %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename(Heroin = Excess_pop)
meth_small <- df_excess_meth %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename(Methamphetamine = Excess_pop)
synth_small <- df_excess_synth %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(state, Excess_pop)) %>% rename("Synthetic Opioids" = Excess_pop)
# We are using a left join to join them
all_types<-Reduce(function(...) merge(..., all=TRUE, by = "state"), list(
  excess_small,
  alc_small,
  benz_small,
  coc_small,
  her_small,
  meth_small,
  synth_small
))
all_types$state = as.character(all_types$state)
all_types <- all_types[order(all_types$state),]

write.csv(all_types,"table.csv", row.names = FALSE)
```

```{r}
excess_small <- df_excess %>% select(c(state, Excess_pop)) %>% rename(Overall = Excess_pop)
alc_small <- df_excess_alc %>% select(c(state, Excess_pop)) %>% rename(Alcohol = Excess_pop)
benz_small <- df_excess_benz %>% select(c(state, Excess_pop)) %>% rename(Benzodiapenes = Excess_pop)
coc_small <- df_excess_coc %>% select(c(state, Excess_pop)) %>% rename(Cocaine = Excess_pop)
her_small <- df_excess_her %>% select(c(state, Excess_pop)) %>% rename(Heroin = Excess_pop)
meth_small <- df_excess_meth %>% select(c(state, Excess_pop)) %>% rename(Methamphetamine = Excess_pop)
synth_small <- df_excess_synth %>% select(c(state, Excess_pop)) %>% rename("Synthetic Opioids" = Excess_pop)
# We are using a left join to join them
all_types<-Reduce(function(...) merge(..., all=TRUE, by = "state"), list(
  excess_small,
  alc_small,
  benz_small,
  coc_small,
  her_small,
  meth_small,
  synth_small
))
all_types$state = as.character(all_types$state)
all_types <- all_types[order(all_types$state),]

write.csv(all_types,"table_numbers.csv", row.names = FALSE)
```


# County Table
```{r}
table_county = df_excess_county %>% mutate (Excess_pop = paste(round(Excess_pop, 1), "(", round(Upper95_pop, 1), "-", round(Lower95_pop, 1), ")")) %>% select(c(County, Excess_pop)) %>% rename("Excess Mortality" = Excess_pop)
write.csv(table_county,"table_county.csv", row.names = FALSE)
```

# Age Analysis

```{r}
library(tidyr)
library(ggpubr)
age_data = read.csv('data/age_data.csv')
age_data = age_data %>% rename(state = Location)
age_data = merge(df_excess, age_data)
age_data = age_data %>% select(state, Excess_pop, Children.0.18, Adults.19.25, Adults.26.34, Adults.35.54, Adults.55.64, X65.)
age_data = gather(age_data, age_group, percent, Children.0.18:X65., factor_key=TRUE)
 # age_data = age_data[age_data$state != 'District of Columbia', ]
```

```{r}
ggplot(age_data, aes(x= percent, y=Excess_pop)) + 
  geom_point() + 
  geom_smooth(method='lm') +
  stat_cor(method = "pearson", label.x.npc = 'left', label.y.npc = 'top') +
  facet_wrap(~ age_group, scales = "free", labeller = labeller(age_group = 
    c("Children.0.18" = "Ages 0-18",
      "Adults.19.25" = "Ages 19-25",
      "Adults.26.34" = "Ages 26-34", 
      "Adults.35.54" = "Ages 35-54",
      "Adults.55.64"= "Ages 55-64", 
      "X65." = "65+")
  )) +
  ylab("Drug Overdose Excess Deaths per 100k") +
  theme_clean()
ggsave(paste('plots/',"Age Scatters"),
  device = "png")
```

```{r}
country_years <- read.delim("data/Country_Year.txt")
  t_county = ts(country_years$Deaths[1:(length(country_years$Deaths) -2)], start = decimal_date(as.Date("2010-01-01")))
  fit <- tryCatch({
        Arima(t_county, order = c(1, 1, 1), method="ML")
    },
    error = function(e){
      Arima(t_county, order = c(2, 2, 2), method="ML")
    })
  f_fit<-forecast(fit, robust = TRUE, h = 2, level = 95)
  true_2020 = country_years$Deaths[(length(country_years$Deaths)-1):(length(country_years$Deaths))]
  true_2020 = ts(true_2020, start = decimal_date(as.Date("2020-01-01")))
```

# Test of Yearly Two-Step Forecast

```{r}
autoplot(t_county, series="Data", size = 0.7) + 
   autolayer(fit$fitted, series= model_name, size = 0.5) +
   autolayer(f_fit, series="Prediction") +
   autolayer(true_2020, series = "True Post-Pandemic") +
  xlab("Year") + ylab("Overdose Deaths") + ggtitle("United States")
```

# Percentage of All Deaths Associated with Drug Types
```{r}
df_percent = matrix(nrow = 80, ncol = 6)
all_death_data <- read.delim("data/all_death_data.txt")
all_death_data = all_death_data %>% select(Month.Code, Deaths)
drug_list = c('Synthetic Opioids', 'Cocaine', 'Meth', 'Benz', 'Alcohol', 'Heroin')
drug_list2 = c('Synthetic Opioids', 'Cocaine', 'Methamphetamines', 'Benzodiazepines', 'Alcohol', 'Heroin')
j = 0
for (i in drug_list){
  j = j + 1
  country_1 <- read.delim(paste0("data/", i, "/Country_Month.txt"))
  country_2 <- read.delim(paste0("data/", i, "/Country_Month_2.txt"))
  country_1 = country_1[country_1$Notes == "", ]
  country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
  country = rbind(country_1, country_2)
  country = country %>% select(Month.Code, Deaths)
  division = as.numeric(country$Deaths)/as.numeric(all_death_data$Deaths)
  df_percent[, j] = division * 100
}

country_1 <- read.delim("data/Year.txt")
country_2 <- read.delim("data/Year_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country = rbind(country_1, country_2)
country = country %>% select(Month.Code, Deaths)
division = as.numeric(country$Deaths)/as.numeric(all_death_data$Deaths)
df_percent <- as.data.frame(df_percent)
df_percent$All = division * 100
colnames(df_percent) <- c(drug_list2, "All Drug Overdoses")
autoplot(ts(df_percent, start = decimal_date(as.Date("2015-01-01")), frequency = 12)) + 
  xlab("Year") + ylab("Percent of All Deaths") + theme_bw()+theme(legend.title = element_blank()) + scale_x_continuous(breaks=c(2015, 2016, 2017, 2018, 2019, 2020, 2021)) + scale_y_continuous(breaks=seq(0, 4, 0.5))
ggsave(paste('plots/',"Percent of All"),
  device = "png", width = 7, height = 5, dpi = 600)
```

```{r}
df_percent = matrix(nrow = 80, ncol = 6)
all_death_data <- read.delim("data/all_death_data.txt")
all_death_data = all_death_data %>% select(Month.Code, Deaths)
drug_list = c('Synthetic Opioids', 'Cocaine', 'Meth', 'Benz', 'Alcohol', 'Heroin')
country_1 <- read.delim("data/Year.txt")
country_2 <- read.delim("data/Year_2.txt")
country_1 = country_1[country_1$Notes == "", ]
country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
country_all_d = rbind(country_1, country_2)
country_all_d = country_all_d %>% select(Month.Code, Deaths)
j = 0
for (i in drug_list){
  j = j + 1
  country_1 <- read.delim(paste0("data/", i, "/Country_Month.txt"))
  country_2 <- read.delim(paste0("data/", i, "/Country_Month_2.txt"))
  country_1 = country_1[country_1$Notes == "", ]
  country_2 = country_2[country_2$Notes == "" & country_2$Month.Code %in% c('2021/01', '2021/02', '2021/03', '2021/04', '2021/05', '2021/06', '2021/07', '2021/08'), ]
  country = rbind(country_1, country_2)
  country = country %>% select(Month.Code, Deaths)
  division = as.numeric(country$Deaths)/as.numeric(country_all_d$Deaths)
  df_percent[, j] = division * 100
}
df_percent <- as.data.frame(df_percent)
colnames(df_percent) <- drug_list2
autoplot(ts(df_percent, start = decimal_date(as.Date("2015-01-01")), frequency = 12)) + 
  xlab("Year") + ylab("Percent of Drug Overdose-Related Deaths") + theme_bw()+theme(legend.title = element_blank()) + scale_x_continuous(breaks=c(2015, 2016, 2017, 2018, 2019, 2020, 2021)) + scale_y_continuous(breaks=seq(10, 70, 10)) 
ggsave(paste('plots/',"Percent of Drugs"),
  device = "png", width = 7, height = 5, dpi = 600)
```
